How Optimal is Word Recognition Under Multimodal Uncertainty?

### 2019-04-12 Writeup/ms.tex
more revisions

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We study how
adults identify novel words under joint uncertainty in the auditory and
visual modalities and we propose an ideal observer model of how cues in
these modalities are combined optimally. Model predictions are tested in
four experiments where recognition is made under various sources of
uncertainty. We found that participants use both auditory and visual
cues to recognize novel words. When the signal is not distorted with
environmental noise, participants weight the auditory and visual cues
optimally, that is, according to the relative reliability of each
modality. In contrast, when one modality has noise added to it, human
perceivers systematically prefer the unperturbed modality to a greater
extent than the optimal model does. This work extends the literature of
perceptual cue combination to the case of word recognition in a
referential context, thus offering a first step towards a model that
accounts for sound-meaning synergies in early word learning.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
probabilistic model that accounts for both types of variability. First,
we describe the model in the simplified case of categorical variability
only. Second, we augment this simplified model to account for sensory
and envirnmental noise.

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. They were paid at an hourly rate
of \$6/hour. Participants were excluded if they reported having
experienced a technical problem of any sort during the online experiment
(N=14), or if they had less than 50\% accurate responses on the
unambiguous training trials (N=6). The final sample consisted of N = 80
participants. All participants provided informed consent before taking
the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

Figure \ref{fig:bimodal} compares the predictions of the normative and
descriptive models against human responses. Remember that the normative
models use the parameters estimated from the unimodal conditions (where
people see input from only one modality) to predict behavior in the
bimodal condition (where people see input from both modalities). The
descriptive model has a similar structure than the optimal model, but is
directly fit to human responses in the bimodal condition in order to
allow us to assess deviation from optimaity.

We found, through comparing the correlation values, that the optimal
model explained more variance than the visual and auditory models did.
However, the optimal model was not perfect: It explained less variance
than the descriptive model did, which indicates a deviation from
optimality. To investigate this deviation, we compare the parameter
values of the optimal model to the values obtaied in the decriptive
model (Table
\ref{tab:param}).\footnote{Note that the descriptive model explained almost all the variance in mean responses, which makes it a reasonble proxy for human real performance in the bimodal condition.}.
We note an increase in \emph{both} the auditory and visual variances.
This increase in noise is compatible with the fact that human responses
appear to be pulled towards chance (i.e., the value O.5) when compared
to the optimal model (see \ref{fig:bimodal}). Below we investigate if
this deviation from optimality can be related to the cue combination
strategy.

@@

This experiment studied the way participants combine multimodal
information to recognize novel words. We found that the optimal model
explained more variance than the auditory or the visual models did,
indicating that particiapants take into account both the auditory and
visual cues when making a decision. That said, Figure \ref{fig:bimodal}
shows that the participants deviated slightly --- but systematically---
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact was captured by the
increase in the value of the variance associated with each modality (as
can be noted from Table \ref{tab:param}). Note, however, that despite
this increase in response randomness, our analysis of modality
preference showed that the \emph{relative} values of these variances
were not different (Figure~\ref{fig:bias}), meaning that there was no
evidence for a modality preference.

@@

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{C_A}+\sigma^2_{N_A} + \sigma^2_{E_A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{C_A}+\sigma^2_{N_A}}.\]

@@

We used the values derived from the unimodal condition to construct the
visual, auditory and optimal models. In addition, we fit a descriptive
model which allowed us to assess real human performance in this
condition. Figure \ref{fig:bimodal} shows that, similar to Experiment 1,
the optimal model explained more variance than the auditory and visual
models did (note, however, that the visual model explained more variance
than the auditory model did). Also similar to Exeperiment 1, the values
of the variances increased in the bimodal condition (Table
\ref{tab:param}).

@@

Here we investigated whethere the observed increase in the auditory and
visual variances affected the relative weighting of the corresponding
modalities. Figure~\ref{fig:bias} (top) shows that the participants'
decision threshold deviated from optimality, and that this deviation was
biased towards the visual modality (the non-noisy modality). Indeed
non-parametric resampling of the data showed a decrease in the value of
the slope in the descriptive model compared to the optimal model
(Figure~\ref{fig:bias}, bottom).

@@

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C}+\sigma^2_{V_N} + + \sigma^2_{V_E}}.\]

@@

Figure \ref{fig:bimodal} shows that almost all the variance was captured
by the auditory model alone, the addition of visual information in the
optomal model did not improve the prediction of human responses. Similar
to Exeperiment 1 and 2, the values of the variances increased in the
bimodal condition (Table \ref{tab:param}).

@@

Experiment 3 tested audi-visual combination in the case where the visual
input was noisy. Whereas in previous experiments the optimal model
explained more variance than the auditory or the visual models did, here
the auditory model alone explained almost all the variance. In other
words, though participants were sensitive to variation in the noisy
visual input when presented in isolation (as shown in Figure
\ref{fig:unimodal}), they tended to ignore this information when the
visual input was presented simultaneusel with the auditory input (i.e.,
in the bimodal condition). Instead, they relied almost exclusively on
the non-noisy auditory
modality.\footnote{The reason why we saw this (floor) effect when we added noise to the visual modality (Experiment 3), and not when we added noise to the auditory modality (Experiment 2), is the fact that our visual stimuli were originally perceived less categorically and with less certainty than the auditory stimuli (see Experiment 1 in \\ref{fig:unimodal}). This fact made it more likely for the visual categorization function to become flat and uninformative after a few drops in precision due to noise on the one had, and to the additional randomness induced by the bimodal presentation on the other hand.}

@@

We recruited a planned sample of \(N=\) 50 participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. They were paid at
an hourly rate of \$7/hour. Participants were excluded if they reported
having experienced a technical problem of any sort during the online
experiment (\(N=\) 0), or if they had less than 75\% accurate responses
on the unambiguous training trials (\(N=\) 7). The final sample
consisted of \(N =\) 43 participants. All participants provided informed
consent before taking the
experiment.\footnote{The sample size, exclusion criteria and the main analysese were pre-registered at https://osf.io/h7mzp/.}

@@

The design and procedure were similar to Experiment 1. There were,
however, two differences: 1) We increased the number of responses
elicited per subject from 70 to 300, and 2) we randomized the order of
the three blocks (i.e., visual-only, auditory-only, and audio-visual)
\emph{within} subject: Each participant saw the 3 blocks exactly 6
times, covering all possible ordering combinations. Unlike the
between-subjet randomzation that we used in Experiment 1, this choices
allowed us to avoid a possible confound linked to the order of exposure.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/weights-1} \caption{The histogram shows the distribution of the participants' predicted (i.e., optimal) cue weighing relative the observed (i.e., descriptive) weigting (see Figure 3 for the details). The density plot shows the distriubtion of simulated data sampled from the population-level probabilistic model. The dashed lines represent 95\% confidence intreval on this simulated distribution. Optimal beahvior is observed when the value of the relative cue weighing is 1 (red solid line). Participants whose values are outside the confidence interval of the simulated dsitribution over-rely on the visual modalirty (left side) or auditory modality (right side) in a non-random way (i.e., with p < 0.05).}\label{fig:weights}
\end{figure}

@@

\subsubsection{Cue combinaton}\label{cue-combinaton}

We analyzed the cue combination strategies at the individual level. For
each participant, we computed the optimal weighing,
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\), relative to the observed (i.e.,
descriptive) weighing, \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). We show
the resulting distribution in Figure \ref{fig:weights}. We note first
that the distribution has a rather bimodal shape, centered around the
optimal cue combination strategy. This finding rules out the hypotheis
that optimality at the population level is a spurious finding, i.e.,
only obtained via aggregating over various sub-otimal strategies.

In addition, we asked whether the observed variance in the individual
distribution was due to the randomness inherent to the process of
sampling from a probabilsitic model or whether it corresponded to a real
between-subject variability induced by different cue weighing
strategies. We simulated responses through sampling from the
population-level models and we computed the resuting distribution of cue
weighing for each simulated indivudual leading to the density function
in Figure \ref{fig:weights}. We can observe that most empirical values
fall within the 95\% confidence interval of the simulated density,
showing that this part of the variance can be due to mere sampling
randomness. However, a few participants had values outide this interval,
indicating that they systematically over-relied on the visual modality
(\(N=\) 5) or the auditory modality (\(N=\) 6) with p-value \$
\textless{} 0.05\$.

@@

Despite this overall near-optomal behavior, we documented two major
cases of sub-optimality. First, in all experiments, the variance
associated with each modality increased in the bimodal condition
compared to the unimodal conditions: Participants responded slightly
more randomly. This increase in randomness could be due limitation on
cognitive resources: Retaining two separate cues at the same time
instead of one cue (as in the unimodal case) is likely to place extra
demands on working memory, causing general performance to drop (see
Mattys \& Wiget, 2011).

Previous research bas found similar cases of suboptimal behavior. For
instance, studies that have explored the identification of ambiguous,
newly learned pairs of word-referent associations have reported what
appears to be a decrease in speech perception acuity in both children
(Stager \& Werker, 1997) and adults (Pajak, Creel, \& Levy, 2016).
Recently, Hofer and Levy (2017) provided a probabilistic model of this
phenomenon. In agreement with the findings of our study, Hofer and Levy
(2017) characterized the apparent reduction in perceptual acuity as an
increase in the noise variance of the auditory modality. Our findings,
besides providing more evidence to this documented fact, suggest that
the reduction in perceptual acuity may occur simultaneously in both the
auditory \emph{and} the visual modalities.

@@

This second case of sub-optimal behavior may be related to the fact that
language understanding under degraded conditions is cognitively more
taxing than language understanding under normal conditions (Mattys et
al., 2012; Peelle, 2018; Rönnberg, Rudner, Lunner, Zekveld, \& others,
2010). Perhaps these demands lead to sub-optimal behavior (i.e.,
over-reliance on the less noisy cue) as participants seek to minimized
cognitive effort. One could also explain this phenomenon in terms of the
metacognitive experience about the fluency with which information is
processed. The perceived perceptual fluency (e.g., the ease with which a
stimulus' physical identity can be identified) can affect a wide variety
of human judgements (see Schwarz, 2004 for a review). In particular,
variables that improve fluency tend to increase liking/preference
(Reber, Winkielman, \& Schwarz, 1998). In our case, the subjective
experience of lower fluency in the noisy modality might cause people to
underestimate information that can be extracted from this modality,
especially when presented simultaneously with a higher fluency
alternative.

@@

Concerning the visual dimension, simulating meaningful variability has
been a notoriousely difficult problem. Following previous studies
(Freedman et al., 2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004),
we used a visual continuum along a one-dimensional morph. This
simplification was motivated by the need to construct a multimodal input
where the auditory and visual components are parametrized in a
symmetrical fashion, allowing us to compare graded effects of auditory
and visual information on categorical judgment. Though such a visual
variability is clearly artificial (one does not encounters in real life
an animal that is, e.g., 30 \% dog and 70 \% cat), we assume that the
induced uncertainty form this visual stimuli has a similar effect on
word recognition as the uncertainty induced by more naturalistic
semantic variability.

@@

Finally, though we used the terminology of word recognition, our work is
only indirectly related to the literature about how a \emph{rich}
lexicon is accessed (e.g., McClelland \& Elman, 1986). We have used this
term in a more specific way, describing access to a simplified lexicon
made of two novel, ambiguous words. Such a simplified experimental
context is not new and has been crucial to our understanding of early
word learning and recognition (e.g., Stager \& Werker, 1997).

@@

Our work provides a formal framework where old and new questions about
word recognition can be given a precise formulation. This framework
enabled us not only to test optimality, but also to examine
systamtically how and by how much people deviate from optimality in
their combination strategies. Thus, our work is part of a growing effort
to go byond optimality tests ---which have limited explanatory power---
and use models that also allow us to identify and explain various
patterns of sub-otimlaity in human behavior (Rahnev \& Denison, 2018).

While we focused on the case of arbitrary associations in novel word
recognition, it is possible to use the same framework to study, for
instance, the case of \textit{iconicity}, that is, when there is a
resemblance between the sound of a word and its referent. Previous work
has suggested that iconicity, among other things, helps with learning
(and generalizing the meaning of) new words (see Dingemanse, Blasi,
Lupyan, Christiansen, \& Monaghan, 2015 for a review). Using the
research strategy in this paper, we can, for example, test whether
iconicity has such an advantage because it mitigates the sub-optimal
patterns observed with more arbitrary pairings.

Finally, though the current framework only characterizes adult word
recognition, it provides a first step towards a model where
developmental questions can also be investigated. For instance, future
work should explore whether children, like adults, use probabilistic
cues from both the auditory and the visual input to recognize ambiguous
words, the extent to which they combine these cues in an optimal
fashion, and whether this cue combination help them to refine their
early phonological and semantic representations.

>>>>

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We study how
adults identify novel words under joint uncertainty in the auditory and
visual modalities and we propose an ideal observer model of how cues in
these modalities are combined optimally. Model predictions are tested in
four experiments where recognition is made under various sources of
uncertainty. We found that participants use both auditory and visual
cues to recognize novel words. When the signal is not distorted with
environmental noise, participants weight the auditory and visual cues
optimally, that is, according to the relative reliability of each
modality. In contrast, when one modality has noise added to it, human
perceivers systematically prefer the unperturbed modality to a greater
extent than the optimal model does. This work extends the literature on
perceptual cue combination to the case of word recognition in a
referential context. In addition, this context offers a link to the
study of multimodal information in word meaning learning.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
probabilistic model that accounts for both types of variability. First,
we describe the model in the simplified case of categorical variability
only. Second, we augment this simplified model to account for sensory
and environmental noise.

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. Participants were excluded if
they reported having experienced a technical problem of any sort during
the online experiment (N=14), or if they had less than 50\% accurate
responses on the unambiguous training trials (N=6). The final sample
consisted of N = 80 participants. All participants provided informed
consent before taking the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

Figure \ref{fig:bimodal} compares the predictions of the normative and
descriptive models against human responses. Remember that the normative
models use the parameters estimated from the unimodal conditions (where
people see input from only one modality) to predict behavior in the
bimodal condition (where people see input from both modalities). The
descriptive model has a similar structure than the optimal model, but is
directly fit to human responses in the bimodal condition in order to
allow us to assess deviation from optimality.

We found, through comparing the correlation values, that the optimal
model explained more variance than the visual and auditory models did.
However, the optimal model was not perfect: It explained less variance
than the descriptive model did, which indicates a deviation from
optimality. To investigate this deviation, we compare the parameter
values of the optimal model to the values obtained in the descriptive
model (Table
\ref{tab:param}).\footnote{Note that the descriptive model explained almost all the variance in mean responses, which makes it a reasonable proxy for human real performance in the bimodal condition.}.
We note an increase in \emph{both} the auditory and visual variances.
This increase in noise is compatible with the fact that human responses
appear to be pulled towards chance (i.e., the value 0.5) when compared
to the optimal model (see \ref{fig:bimodal}). Below we investigate if
this deviation from optimality can be related to the cue combination
strategy.

@@

This experiment studied the way participants combine multimodal
information to recognize novel words. We found that the optimal model
explained more variance than the auditory or the visual models did,
indicating that participants take into account both the auditory and
visual cues when making a decision. That said, Figure \ref{fig:bimodal}
shows that the participants deviated slightly --- but systematically---
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact was captured by the
increase in the value of the variance associated with each modality (as
can be noted from Table \ref{tab:param}). Note, however, that despite
this increase in response randomness, our analysis of modality
preference showed that the \emph{relative} values of these variances
were not different (Figure~\ref{fig:bias}), meaning that there was no
evidence for a modality preference.

@@

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N} + \sigma^2_{A_E}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C}+\sigma^2_{V_N}}.\]

@@

We used the values derived from the unimodal condition to construct the
visual, auditory and optimal models. In addition, we fit a descriptive
model which allowed us to assess real human performance in this
condition. Figure \ref{fig:bimodal} shows that, similar to Experiment 1,
the optimal model explained more variance than the auditory and visual
models did (note, however, that the visual model explained more variance
than the auditory model did). Also similar to Experiment 1, the values
of the variances increased in the bimodal condition (Table
\ref{tab:param}).

@@

Here we investigated whether the observed increase in the auditory and
visual variances affected the relative weighting of the corresponding
modalities. Figure~\ref{fig:bias} (top) shows that the participants'
decision threshold deviated from optimality, and that this deviation was
biased towards the visual modality (the non-noisy modality). Indeed
non-parametric resampling of the data showed a decrease in the value of
the slope in the descriptive model compared to the optimal model
(Figure~\ref{fig:bias}, bottom).

@@

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C}+\sigma^2_{V_N} + \sigma^2_{V_E}}.\]

@@

Figure \ref{fig:bimodal} shows that almost all the variance was captured
by the auditory model alone, the addition of visual information in the
optimal model did not improve the prediction of human responses. Similar
to Experiments 1 and 2, the values of the variances increased in the
bimodal condition (Table \ref{tab:param}).

@@

Experiment 3 tested audi-visual combination in the case where the visual
input was noisy. Whereas in previous experiments the optimal model
explained more variance than the auditory or the visual models did, here
the auditory model alone explained almost all the variance. In other
words, though participants were sensitive to variation in the noisy
visual input when presented in isolation (as shown in Figure
\ref{fig:unimodal}), they tended to ignore this information when the
visual input was presented simultaneously with the auditory input (i.e.,
in the bimodal condition). Instead, they relied almost exclusively on
the non-noisy auditory
modality.\footnote{The reason why we saw this (floor) effect when we added noise to the visual modality (Experiment 3), and not when we added noise to the auditory modality (Experiment 2), is the fact that our visual stimuli were originally perceived less categorically and with less certainty than the auditory stimuli (see Experiment 1 in \\ref{fig:unimodal}). This fact made it more likely for the visual categorization function to become flat and uninformative after a few drops in precision due to noise on the one had, and to the additional randomness induced by the bimodal presentation on the other hand.}

@@

We recruited a planned sample of \(N=\) 50 participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. Participants were
excluded if they reported having experienced a technical problem of any
sort during the online experiment (\(N=\) 0), or if they had less than
75\% accurate responses on the unambiguous training trials (\(N=\) 7).
The final sample consisted of \(N =\) 43 participants. All participants
provided informed consent before taking the
experiment.\footnote{The sample size, exclusion criteria and the main analyses were pre-registered at https://osf.io/h7mzp/.}

@@

The design and procedure were similar to Experiment 1. There were,
however, two differences: 1) We increased the number of responses
elicited per subject from 70 to 300, and 2) we randomized the order of
the three blocks (i.e., visual-only, auditory-only, and audio-visual)
\emph{within} subject: Each participant saw the 3 blocks exactly 6
times, covering all possible ordering combinations. Unlike the
between-subject randomization that we used in Experiment 1, this choices
allowed us to avoid a possible confound linked to the order of exposure.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/weights-1} \caption{The histogram shows the distribution of the participants' predicted (i.e., optimal) cue weighting relative the observed (i.e., descriptive) weighting (see Figure 3 for the details). The density plot shows the distribution of simulated data sampled from the population-level probabilistic model. The dashed lines represent 95\% confidence interval on this simulated distribution. Optimal behavior is observed when the value of the relative cue weighting is 1 (red solid line). Participants whose values are outside the confidence interval of the simulated distribution over-rely on the visual modality (left side) or auditory modality (right side) beyond sampling-related randomness (i.e., with p < 0.05).}\label{fig:weights}
\end{figure}

@@

\subsubsection{Cue combination}\label{cue-combination-3}

We analyzed the cue combination strategies at the individual level. For
each participant, we computed the optimal weighting,
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\), relative to the observed (i.e.,
descriptive) weighting, \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). We show
the resulting distribution in Figure \ref{fig:weights}. We note first
that the distribution has a rather bimodal shape, centered around the
optimal cue combination strategy. This finding rules out the hypothesis
that optimality at the population level is a spurious finding, i.e.,
only obtained by aggregating over various sub-optimal strategies.

In addition, we asked whether the observed variance in the individual
distribution was due to the randomness inherent to the process of
sampling from a probabilistic model or whether it corresponded to a real
between-subject variability induced by different cue weighting
strategies. We simulated responses through sampling from the
population-level models and we computed the resulting distribution of
cue weighting for each simulated individual leading to the density
function in Figure \ref{fig:weights}. We can observe that most empirical
values fall within the 95\% confidence interval of the simulated
density, showing that this part of the variance can be due to mere
sampling randomness. However, a few participants had values outside this
interval, indicating that they systematically over-relied on the visual
modality (\(N=\) 5, \(p < 0.05\)) or the auditory modality (\(N=\) 6,
\(p < 0.05\)).

@@

Despite this overall near-optimal behavior, we documented two major
cases of sub-optimality. First, in all experiments, the variance
associated with each modality increased in the bimodal condition
compared to the unimodal conditions: Participants responded slightly
more randomly. This increase in randomness could be due limitation on
cognitive resources: Processing two separate --- and perceptually
uncorrelated --- cues instead of one cue (as in the unimodal case) is
likely to place extra demands on working memory, causing general
performance to drop (see Mattys \& Wiget, 2011).

Previous research has found similar cases of suboptimal behavior. For
instance, studies that have explored the identification of ambiguous,
newly learned pairs of word-referent associations have reported what
appears to be a decrease in speech perception acuity in both children
(Stager \& Werker, 1997) and adults (Pajak, Creel, \& Levy, 2016). In
agreement with the findings of our study, Hofer and Levy (2017)
characterized the apparent reduction in perceptual acuity as an increase
in the noise variance of the auditory modality. Our findings, besides
providing more evidence to this documented fact, suggest that the
reduction in perceptual acuity may occur simultaneously in both the
auditory \emph{and} the visual modalities.

@@

This second case of sub-optimal behavior may be related to the fact that
language understanding under degraded conditions is cognitively more
taxing than language understanding under normal conditions (Mattys et
al., 2012; Peelle, 2018; Rönnberg, Rudner, Lunner, Zekveld, \& others,
2010). Perhaps these demands lead to sub-optimal behavior (i.e.,
over-reliance on the less noisy cue) as participants seek to minimize
cognitive effort. One could also explain this phenomenon in terms of the
metacognitive experience about the fluency with which information is
processed. The perceived perceptual fluency (e.g., the ease with which a
stimulus' physical identity can be identified) can affect a wide variety
of human judgements (see Schwarz, 2004 for a review). In particular,
variables that improve fluency tend to increase liking/preference
(Reber, Winkielman, \& Schwarz, 1998). In our case, the subjective
experience of lower fluency in the noisy modality might cause people to
underestimate information that can be extracted from this modality,
especially when presented simultaneously with a higher fluency
alternative.

@@

Concerning the visual dimension, simulating meaningful variability has
been a notoriously difficult problem. Following previous studies
(Freedman et al., 2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004),
we used a visual continuum along a one-dimensional morph. This
simplification was motivated by the need to construct a multimodal input
where the auditory and visual components are parametrized in a
symmetrical fashion, allowing us to compare graded effects of auditory
and visual information on categorical judgment. Though such a visual
variability is clearly artificial (one does not encounters in real life
an animal that is, e.g., 30 \% dog and 70 \% cat), we assume that the
induced uncertainty form this visual stimuli has a similar effect on
word recognition as the uncertainty induced by more naturalistic
semantic variability.

@@

Finally, though we used the terminology of word recognition, our work is
only indirectly related to the literature about how a rich lexicon is
accessed (e.g., McClelland \& Elman, 1986). We have used this term in a
more specific way, describing access to a simplified lexicon made of two
novel, ambiguous words. Such a simplified experimental context is not
new and has been crucial to our understanding of early word learning and
recognition (e.g., Stager \& Werker, 1997).

@@

Our work used an ideal observer model to study word recognition under
audio-visual uncertainty. This framework enabled us not only to test
optimality, but also to examine systematically how and by how much
people deviate from optimality in their combination strategies. Thus,
our work is part of a growing effort to go beyond optimality tests ---
which have limited explanatory power --- and use models that also allow
us to identify and explain various patterns of sub-optimality in human
behavior (Rahnev \& Denison, 2018).

While we focused on the case of arbitrary associations in novel word
recognition, it is possible to extend the framework to other cases such
as that of \textit{iconicity}, i.e., when there is a resemblance between
the sound of a word and its referent. Previous work has suggested that
iconicity, among other things, helps with learning (and generalizing the
meaning of) new words (see Dingemanse, Blasi, Lupyan, Christiansen, \&
Monaghan, 2015 for a review). Using the research strategy in this paper,
we can, for example, test whether iconicity has such an advantage
because it mitigates the sub-optimal patterns observed with more
arbitrary pairings.

Finally, though the current framework only characterizes adult novel
word recognition, it provides a first step towards a model where
developmental questions can also be investigated. For instance, future
work should explore whether children, like adults, use probabilistic
cues from both the auditory and the visual input to recognize ambiguous
words, the extent to which they combine these cues in an optimal
fashion, and whether this cue combination help them learn words and
refine their early phonological and semantic representations.
